---
title: Quantitative approach to topology generators
authors:
  - name: Tanya Strydom
    id: ts
    orcid: 0000-0001-6067-1349
date: last-modified
bibliography: ../references.bib
---

I think in the long run I will probably turn the other repo ([BecksLab/topology_generators](https://github.com/BecksLab/topology_generators)) into a project that we can then access and I will port all of the working code here... I think it will just come down to what the 'cost' is of the computational side of things and how convoluted it might end up looking.

Anyway so for now I am just pulling in the INTERIM data from some of the models. This uses (for now) 20 mangal [@poisotMangalMakingEcological2016] datasets and just generates some networks for using the niche and cascade models. The 'results' for now is just counting the number of links each model generates for each network because that's what my brain decided on doing...

```{r}
#| include: false
#| warning: false
library(tidyverse)

# download file from other repo
download.file("https://raw.githubusercontent.com/BecksLab/topology_generators/main/data/topology_summary.csv", "../notebooks/data/topology_models.csv")

df = read.csv("data/topology_models.csv")  %>% 
        filter(complexity_real != 0) %>%
        pivot_longer(
            cols = c(complexity_real, distance_real, connectance_real, basal_real, top_real), 
            names_to = "real",
            values_to = "real_val")  %>% 
        pivot_longer(
            cols = c(complexity_mod, distance_mod, connectance_mod, basal_mod, top_mod), 
            names_to = "test",
            values_to = "model_val")  %>% 
        mutate(
            real = str_extract(real, "[^_]*"),
            test = str_extract(test, "[^_]*"))  %>% 
        filter(real == test)
```

Lets tart by summarising the 

```{r}
#| warning: false
#| echo: false
#| label: fig-boxplot
#| fig-cap: "Boxplot looking at raw values for each measurement for each model"

ggplot(df) +
    geom_boxplot(aes(x = real,
                    y = real_val),
                colour = 'black') +
    geom_boxplot(aes(x = real,
                    y = model_val,
                    colour = model)) +
    facet_wrap(vars(real),
                scales = 'free') +
    scale_size(guide = 'none') +
    coord_cartesian(expand = FALSE) +
    theme_classic() +
    theme(panel.border = element_rect(colour = 'black',
                                      fill = "#ffffff00"))

```


Now we can look at the Z scores for the different models for the different network measures that we use

```{r}
#| warning: false
#| echo: false
#| label: fig-topology
#| fig-cap: "Z-scores for network summary statistics. Negative Z-scores indicate a (mean) value greater than expected. The magnitude of Z-score probably also tells us how 'variable'/constrained the model is... Note here that 'basal' is calculated as the proportion of species that have a generality value of zero *i.e.,* are basal, for 'top' species it is the same except vulnerability is used. I am aware that there are no Z scores for some of the Random models and am looking into it"

ggplot(df %>% 
        group_by(id, real, model)  %>% 
        reframe(x_real = real_val,
                mu_sim = mean(model_val, na.rm = TRUE),
                sd_sim = sd(model_val, na.rm = TRUE)) %>%
        mutate(z_score = ((x_real-mu_sim)/sd_sim)) %>%
        distinct()) +
    geom_vline(aes(xintercept = 0)) +
    geom_histogram(aes(x = z_score,
                    fill = model),
                colour = "#ffffff00") +
    facet_grid(rows = vars(model),
                cols = vars(real),
                scales = "free") +
    scale_size(guide = 'none') +
    coord_cartesian(expand = FALSE) +
    theme_classic() +
    theme(panel.border = element_rect(colour = 'black',
                                      fill = "#ffffff00"))

```

## References {.unnumbered}

::: {#refs}
:::