---
title: Qualitative approach to topology generators
author: Tanya Strydom
bibliography: ../references.bib
---

## Venn Diagram

Here we will use the standard [REF FOR TIDYVERSE] and @gaoGgVennDiagramIntuitiveEasytoUse2021 to create a venn diagram for the quantitative portion of or generator/model overview process. Also remember to drop the R version (maybe also look into if you can `renv` this because we aim for reproducibility)

```{r}
#| include: false
library(tidyverse)
library(ggVennDiagram)
```


First we need to create the df with the relevant 'traits' for each model

```{r}
#| include: false

x <- list(Niche=c("species agnostic", "structural", "synthetic network"),
          ADBM=c("node-level", "mechanistic"),
          `Transfer`=c("node-level", "embedding", "metaweb"),
          PFIM=c("node-level", "mechanistic", "metaweb"),
          Cascade=c("species agnostic", "structural", "synthetic network"))

venn <- Venn(x)
data <- process_data(venn)
```

Create venn diagram:

```{r}
#| echo: false
#| label: fig-venn
#| fig-cap: "Venn diagram for qualitative analysis/overview of the fancy maths things"
#| fig-alt: "TODO"
ggVennDiagram(x, 
              label = "none") +
  theme_void() +
  theme(legend.position = 'none')
```

Looking at it now I don't think this is the wayTM and trying to sort of 'ordinate' modelling space might actually be more meaningful... Specifically thinking here that I need to also actually just read about model comparison (but without the performance aspect)

## Ordination approach

So at least according to @mccuneAnalysisEcologicalCommunities2002 we should be able to ordinate using only categorical data although we should stick to an NMDS. Its just going to be a case of encoding said categories correctly. How I see it is that each category that we 'create' will be its own entry (we could maybe also just turn each category into a factor but this comes at the cost of assuming the distribution of said variable, well that's what I infer form my 2 minute skim read).

## References {.unnumbered}

::: {#refs}
:::